---
title: "dummy"
author: "Khoi"
date: "2023-06-08"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

First, we create the custom function `BasicSummary` to summarize preliminary specifications of the variables/columns in a data frame. The function has 2 arguments: a data frame `df` and decimal values precision `dgts`, defaulted at 2.

```{r basic summary}
BasicSummary <- function(df, dgts=2){

  n <- nrow(df)
  m <- ncol(df)
  varNames  <- colnames(df);          topCount  <- vector("numeric",m)
  varType   <- vector("character",m); missCount <- vector("numeric",m)
  topLevel  <- vector("character",m); levels    <- vector("numeric",m)
  
  for (i in 1:m){
    x <- df[,i]
    xtab <- table(x, useNA="ifany")
    nums <- as.numeric(xtab)
    lvls <- names(xtab)
    
    varType[i]   <- class(x)
    topLevel[i]  <- lvls[which.max(nums)]
    topCount[i]  <- max(nums)
    missCount[i] <- length(which((is.na(x)) | (x=="") | (x==" ")))
    levels[i]    <- length(xtab)
  }
  
  topFrac  <- round(topCount/n*100,  digits=dgts)
  missFrac <- round(missCount/n*100, digits=dgts)
  summaryFrame <- data.frame(
    variable = varNames,  dtype    = varType,
    levels   = levels,    topLevel = topLevel,
    topCount = topCount,  topFrac  = topFrac,
    missFreq = missCount, missFrac = missFrac)
  return(summaryFrame)
}
```

The function returns a data frame with one row for each column of `df` and the following columns:

-   `variable`: variable/column name, or properties of each data point.

-   `dtype`: data type/class of the variable.

-   `levels`: number of distinct values in a variable. The standard missing value `NA` is also considered as a level.

-   `topLevel`: most frequently occurring value.

-   `topCount`: number of times `topLevel` occurs.

-   `topPct`: percentage `topLevel` occurs compared to the whole data set `df`.

-   `missFreq`: number of missing value in the variable. The 3 types are `NA`, blanks, and spaces.

-   `missPct`: percentage `missFreq` occurs compared to the whole data set `df`.

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
